---
title: "Package Functions (Week 3)"
output:
  html_notebook:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

# Preliminary Functions (format, maybe warnings and so)

# Transformation functions (ilr, inverseolr,...)

## Function that creates the Helmert Matrix

If $V$ (contrast matrix) is not passed, computations are done with the Helmert Matrix as default.

```{r} 
# Creates de Helmert Matrix given the number of parts of the variable
V_D = function(D){ # D is the dimension (parts)
  V_D = matrix(0L, nrow =D , ncol = D-1)
  for (i in 1:(D-1)){  # pseudodiagonal elements
    V_D[i,i] = (D-i)/(sqrt((D-i+1)*(D-i)))
  }
  for (i in 2:D){ # below pseudodiagonal elements
    for (j in 1:i-1){
      V_D[i,j]=-1/(sqrt((D-j+1)*(D-j)))
    }
  } 
  return(V_D)
}
```

```{r}
V_D(3)
```



```{r}
ilr = function(x){ # x matrix N x D
  x = C(x)
  V = V_D(dim(x)[2]) 
  return(t(t(V)%*%t(log(x))))
}
```



# Functions for the Regression

Assume we have $k$ explanatory variables (some compositional and some not) and $N$ observations. The dependent variable $Y$ can be compositional or not. In this first version there are four inputs: $Y$, the independent variable which can be compositional (comp=1) or not (comp=0), X is the list of non compositional variables and Z the list of compositional variables. Each variable inside the list is a matrix nxD (D≥1 the number of parts).

Recall that the regression in the ILR coordinate space is defined by:

$\mathrm{ilr}(\mathbf{Y}_i) = \mathbf{b}_0^*+ \sum_{q=1}^Q\mathrm{ilr}(\mathbf{X}_{qi})\mathbf{B}^*_q + \sum_{k=1}^K Z_{ki} \mathbf{b}^*_k + \mathrm{ilr}(\epsilon_i)$


```{r}
  c_lm = function(Y,X,Z,comp){ 
    if (missing(X)){
      X=list(matrix())
    }
    if (missing(Z)){
      Z=list(matrix())
    }
    if (comp==1){
      ty = ilr(Y)
    }
    else{
      ty = Y
    }
    tx = matrix(NaN,dim(Y)[1])
    if (!is.na(X[[1]][1,1])){
      for (var in X){
        tx = cbind(tx,ilr(var)) ## here compositions::ilr to recover their result
      }
    }
    if (!is.na(Z[[1]][1,1])){
      for (var in Z){
        tx = cbind(tx,var)
      }
    }
    tx = tx[,-1]
    print("Transformed endogenous:")
    print(ty)
    print("--------------------------")
    print("Transformed exogenous")
    print(tx)
    print("--------------------------")
    
    res = mlm(ty,tx)
    print("B matrix in the ILR space:")
    print(res)
    print("--------------------------")
    print("B in the simplex:")
    print(inverseilr(t(inverseilr(t(res)))))
  }
```


# Functions on Impacts